{\bf {\Large Improving estimates of ecosystem metabolism computed from dissolved oxygen time series}}

### 2014-11-04 swmp_noise.tex
some edits from Jim's comments, windows specific to each case study

The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from 0.09 (as a proportion of the total estimates, Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.13 for production, 0.13 for respiration), and only slightly reduced for Elkhorn Slough (0.17 for production, 0.17 for respiration).  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-55.5 \% change from the annual mean) and respiration (-55.2 \%) for Elkhorn Slough, increased mean production (74.0 \%) and respiration (74.8 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Mean net ecosystem metabolism was unchanged for all sites.  Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after filtering.

@@

% plots of summarized metabolism estimates, before/after detiding
% ELKVM, PDBBY
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/metab_sum1}

@@

\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.  Aggregated estimates are for Elkhorn Slough and Padilla Bay from observed and filtered \ac{DO} time series.\label{fig:metab_sum1}}
\end{figure}

% RKBMB, SAPDC
% plots of summarized metabolism estimates, before/after detiding
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/metab_sum2}

\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month and season.  Aggregated estimates are for Rookery Bay and Sapelo Island from observed and filtered \ac{DO} time series.  May was removed from Rookery Bay because of incomplete data.\label{fig:metab_sum2}}
\end{figure}

@@

% summary of simulation performance for detided and biological, sim parameters
% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1") 
%
\begin{table}[h]
\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).  Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which the diel \ac{DO} component was zero).\label{tab:dtd_perf1}} 
\begin{center}
\begin{tabular}{llllllclllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
\cline{2-6} \cline{8-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.30&0.51&0.82&1.00&&0.00&0.68&1.10&1.97&2.39\tabularnewline
~~1&-0.28&0.38&0.59&0.88&1.00&&0.00&0.59&1.07&1.96&2.40\tabularnewline
~~2&-0.39&0.46&0.63&0.90&1.00&&0.00&0.62&1.10&1.97&2.40\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
~~0& 0.00&0.27&0.58&0.93&1.00&&0.00&0.34&1.00&1.96&2.12\tabularnewline
~~1&-0.78&0.37&0.58&0.83&1.00&&0.00&0.63&1.09&1.98&2.12\tabularnewline
~~2&-0.78&0.47&0.61&0.82&1.00&&0.00&0.98&1.34&1.99&2.40\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.34&0.57&0.86&1.00&&0.00&0.63&1.06&1.96&2.40\tabularnewline
~~1&-0.78&0.37&0.59&0.85&1.00&&0.00&0.63&1.06&1.97&2.40\tabularnewline
~~2&-0.78&0.41&0.61&0.85&1.00&&0.00&0.63&1.11&1.98&2.40\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.31&0.82&0.98&1.00&&0.00&0.29&0.76&1.50&2.40\tabularnewline
~~1& 0.05&0.37&0.58&0.81&0.99&&0.07&0.98&1.05&1.49&2.39\tabularnewline
~~2& 0.05&0.40&0.58&0.70&0.99&&0.15&1.06&1.96&2.01&2.40\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

% summary of simulation performance for detided and biological, window widths
% latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),      cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,      colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",          "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2") 
%
\begin{table}[!tbp]
\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).  Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}} 
\begin{center}
\begin{tabular}{llllllclllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Window}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
\cline{2-6} \cline{8-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
\hline
{\bfseries Days}&&&&&&&&&&&\tabularnewline
~~1&-0.78&0.63&0.89&0.97&1.00&&0.00&0.28&0.59&1.04&2.12\tabularnewline
~~3&-0.07&0.40&0.59&0.75&1.00&&0.00&0.99&1.08&1.28&2.08\tabularnewline
~~6& 0.00&0.26&0.40&0.58&1.00&&0.00&1.95&1.98&2.05&2.40\tabularnewline
\hline
{\bfseries Hours}&&&&&&&&&&&\tabularnewline
~~1&-0.78&0.36&0.58&0.82&1.00&&0.00&0.63&1.11&1.96&2.40\tabularnewline
~~3& 0.00&0.40&0.60&0.87&1.00&&0.00&0.58&1.07&1.97&2.36\tabularnewline
~~6& 0.03&0.37&0.59&0.85&1.00&&0.00&0.64&1.10&1.98&2.40\tabularnewline
\hline
{\bfseries Tide}&&&&&&&&&&&\tabularnewline
~~0.25& 0.00&0.42&0.63&0.91&1.00&&0.00&0.51&1.04&1.97&2.21\tabularnewline
~~0.5& 0.06&0.43&0.62&0.88&1.00&&0.00&0.61&1.09&1.97&2.27\tabularnewline
~~1&-0.78&0.30&0.51&0.79&1.00&&0.00&0.73&1.20&1.97&2.40\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

% descriptive table of case studies
% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",          "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),      n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",          "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",          "NEM"), label = "tab:case_att") 
%
\begin{table}[!tbp]
\caption{Summary statistics of tidal component amplitudes (m), selected water quality parameters (\ac{DO} mg L$^{-1}$, chlorophyll-a $\mu$g L$^{-1}$, salinity psu, water temperature $^{\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.  Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).  Water quality data are averages for the entire period of record for each site.  Metabolism estimates are means of daily integrated values.\label{tab:case_att}} 
\begin{center}
\begin{tabular}{lllllcllllclll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{4}{c}{\bfseries Tidal amplitude}&\multicolumn{1}{c}{\bfseries }&\multicolumn{4}{c}{\bfseries Water quality}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Metabolism\textsuperscript{\textit{a}}}\tabularnewline
\cline{2-5} \cline{7-10} \cline{12-14}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{O1}&\multicolumn{1}{c}{P1}&\multicolumn{1}{c}{M2}&\multicolumn{1}{c}{S2}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Chl}&\multicolumn{1}{c}{Sal}&\multicolumn{1}{c}{Temp}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Pg}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
ELKVM&0.24&0.12&0.48&0.13&&7.87&3.87&32.43&13.78&&8.14&-8.19&-0.05\tabularnewline
PDBBY&0.46&0.23&0.63&0.15&&8.97&2.24&29.17&10.44&&5.95&-5.90& 0.05\tabularnewline
RKBMB&0.13&0.04&0.36&0.10&&4.48&4.50&30.53&25.85&&3.02&-3.62&-0.60\tabularnewline
SAPDC&0.10&0.02&0.54&0.07&&4.96&5.98&27.30&21.77&&4.89&-6.04&-1.16\tabularnewline
\hline
\end{tabular}
\end{center}
\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

% correlations with tide before/after wtreg
% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Filtered"), 4), label = "tab:cor_res") 
%
\begin{table}[!tbp]
\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (filtered) filtering with weighted regression.  \ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.\label{tab:cor_res}} 
\begin{center}
\begin{tabular}{lllll}
\hline\hline
\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
{\bfseries ELKVM}&&&&\tabularnewline
~~Observed& 0.47***& 0.60***& 0.73***& 0.35***\tabularnewline
~~Filtered& 0.02*& 0.19***& 0.13*& 0.06 \tabularnewline
\hline
{\bfseries PDBBY}&&&&\tabularnewline
~~Observed&-0.45***&-0.33***&-0.46***&-0.25***\tabularnewline
~~Filtered& 0.07***& 0.48***& 0.47***&-0.21***\tabularnewline
\hline
{\bfseries RKBMB}&&&&\tabularnewline
~~Observed& 0.28***& 0.34***& 0.39***& 0.24***\tabularnewline
~~Filtered&-0.02**&-0.31***&-0.36***& 0.12*\tabularnewline
\hline
{\bfseries SAPDC}&&&&\tabularnewline
~~Observed& 0.48***& 0.54***& 0.71***& 0.41***\tabularnewline
~~Filtered&-0.03***& 0.16**& 0.18***&-0.05 \tabularnewline
\hline
\end{tabular}
\end{center}
\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

>>>>

The proportion of daily integrated metabolism estimates that were anomalous (negative production, positive respiration) were significantly reduced for most sites after filtering (\cref{tab:case_res}), perhaps indicating the relative effects of water movement.  Before filtering, anomalous values ranged from 0.09 (as a proportion of the total estimates, Rookery Bay) to 0.22 (Padilla Bay) for production and 0.08 (Rookery Bay) to 0.21 (Elkhorn Slough) for respiration. Anomalous values were reduced to near zero for Rookery Bay and Sapelo Island, by approximately half for Padilla Bay (0.08, 0.07, 0.06, 0.06, 0.07, 0.08, 0.10, 0.11, 0.07, 0.06, 0.06, 0.08, 0.14, 0.08, 0.09, 0.06, 0.05, 0.05, 0.06, 0.14, 0.06, 0.04, 0.04, 0.06, 0.06, 0.18, 0.08, 0.04, 0.05, 0.06, 0.04, 0.05, 0.06, 0.28, 0.10, 0.06, 0.06, 0.07, 0.06, 0.35, 0.16, 0.08, 0.07, 0.06, 0.05, 0.07, 0.12, 0.08, 0.06, 0.04, 0.07, 0.07, 0.06, 0.06, 0.06, 0.12, 0.05, 0.04, 0.05, 0.07, 0.06, 0.16, 0.07, 0.05, 0.04, 0.05, 0.22, 0.11, 0.05, 0.05, 0.08, 0.06, 0.05, 0.08, 0.13, 0.14, 0.15, 0.06, 0.06, 0.07, 0.07, 0.07, 0.06, 0.11, 0.05, 0.06, 0.06, 0.06, 0.15, 0.08, 0.04, 0.04, 0.04, 0.05, 0.17, 0.10, 0.04, 0.04, 0.06, 0.05, 0.07, 0.10, 0.11, 0.06, 0.14, 0.07, 0.06, 0.08, 0.08, 0.07, 0.08, 0.07, 0.05, 0.05, 0.06, 0.06, 0.14, 0.07, 0.04, 0.04, 0.06, 0.18, 0.09, 0.04, 0.04 for production, 0.07, 0.07, 0.06, 0.06, 0.06, 0.08, 0.10, 0.11, 0.08, 0.06, 0.06, 0.08, 0.19, 0.07, 0.09, 0.06, 0.05, 0.05, 0.06, 0.12, 0.06, 0.04, 0.04, 0.06, 0.06, 0.19, 0.08, 0.04, 0.05, 0.06, 0.04, 0.05, 0.06, 0.33, 0.11, 0.06, 0.05, 0.07, 0.06, 0.39, 0.20, 0.08, 0.07, 0.06, 0.06, 0.07, 0.12, 0.08, 0.06, 0.04, 0.08, 0.07, 0.06, 0.06, 0.06, 0.14, 0.04, 0.04, 0.06, 0.07, 0.07, 0.21, 0.06, 0.04, 0.04, 0.05, 0.34, 0.12, 0.04, 0.04, 0.08, 0.06, 0.06, 0.07, 0.12, 0.13, 0.15, 0.07, 0.07, 0.07, 0.06, 0.07, 0.06, 0.11, 0.06, 0.06, 0.06, 0.06, 0.17, 0.07, 0.04, 0.04, 0.04, 0.06, 0.25, 0.11, 0.04, 0.05, 0.06, 0.06, 0.07, 0.10, 0.10, 0.06, 0.14, 0.07, 0.07, 0.07, 0.09, 0.08, 0.09, 0.07, 0.05, 0.05, 0.07, 0.07, 0.12, 0.08, 0.04, 0.04, 0.06, 0.20, 0.10, 0.04, 0.04 for respiration), and only slightly reduced for Elkhorn Slough (0.18, 0.07, 0.05, 0.22, 0.24, 0.19, 0.18, 0.16, 0.15, 0.21, 0.19, 0.13, 0.18, 0.13, 0.16, 0.15, 0.12, 0.09, 0.05, 0.18, 0.13, 0.10, 0.05, 0.14, 0.03, 0.20, 0.12, 0.10, 0.05, 0.02, 0.10, 0.08, 0.08, 0.26, 0.17, 0.12, 0.09, 0.16, 0.09, 0.33, 0.21, 0.14, 0.13, 0.13, 0.21, 0.20, 0.16, 0.13, 0.15, 0.11, 0.15, 0.15, 0.15, 0.11, 0.11, 0.18, 0.16, 0.12, 0.09, 0.14, 0.08, 0.18, 0.16, 0.12, 0.09, 0.08, 0.23, 0.18, 0.12, 0.08, 0.13, 0.11, 0.22, 0.23, 0.17, 0.14, 0.13, 0.16, 0.18, 0.16, 0.13, 0.16, 0.09, 0.16, 0.16, 0.12, 0.11, 0.08, 0.16, 0.15, 0.12, 0.07, 0.13, 0.07, 0.19, 0.15, 0.12, 0.08, 0.07, 0.22, 0.23, 0.17, 0.18, 0.10, 0.15, 0.15, 0.19, 0.18, 0.14, 0.11, 0.16, 0.17, 0.12, 0.10, 0.07, 0.07, 0.16, 0.15, 0.11, 0.07, 0.05, 0.19, 0.13, 0.11, 0.07 for production, 0.19, 0.05, 0.06, 0.24, 0.23, 0.18, 0.18, 0.16, 0.19, 0.24, 0.21, 0.13, 0.21, 0.10, 0.15, 0.18, 0.13, 0.10, 0.06, 0.19, 0.14, 0.10, 0.06, 0.15, 0.04, 0.23, 0.14, 0.10, 0.06, 0.03, 0.11, 0.08, 0.08, 0.30, 0.18, 0.14, 0.10, 0.15, 0.09, 0.36, 0.22, 0.14, 0.11, 0.11, 0.24, 0.20, 0.16, 0.11, 0.15, 0.11, 0.16, 0.16, 0.16, 0.11, 0.09, 0.17, 0.17, 0.13, 0.09, 0.13, 0.08, 0.23, 0.16, 0.13, 0.09, 0.08, 0.29, 0.18, 0.12, 0.09, 0.13, 0.08, 0.24, 0.23, 0.16, 0.15, 0.13, 0.18, 0.19, 0.16, 0.12, 0.18, 0.09, 0.16, 0.17, 0.13, 0.12, 0.08, 0.20, 0.15, 0.13, 0.08, 0.14, 0.07, 0.27, 0.16, 0.11, 0.09, 0.07, 0.24, 0.23, 0.17, 0.18, 0.09, 0.15, 0.19, 0.22, 0.18, 0.14, 0.10, 0.16, 0.17, 0.13, 0.11, 0.07, 0.07, 0.16, 0.14, 0.11, 0.07, 0.05, 0.25, 0.15, 0.11, 0.07 for respiration).  Metabolism estimates using filtered \ac{DO} time series had decreased mean production (-69.3 \% change from the annual mean) and respiration (-69.3 \%) for Elkhorn Slough, increased mean production (43.7 \%) and respiration (44.7 \%) for Padilla Bay, and generally unchanged mean production and respiration for Rookery Bay and Sapelo Island (\cref{tab:case_res}).  Mean net ecosystem metabolism was unchanged for all sites.  Decreases in the standard erorr for all metabolism estimates (production, respiration, and net) were observed for all cases after filtering.

@@

% plots of summarized metabolism estimates, before/after detiding
% all case studies
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/metab_sum}

@@

\caption[Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month]{Means and standard errors of daily metabolism estimates (gross production, total respiration) aggregated by month.  Aggregated estimates are shown for observed and filtered \ac{DO} time series.  May was removed from Rookery Bay because of incomplete data.\label{fig:metab_sum}}
\end{figure}

@@

% summary of simulation performance for detided and biological, sim parameters
%latex.default(tab, file = "", where = "h", rowlabel = "Parameter",     caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,         4), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,         5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",         "Median", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1")%
\begin{table}[h]
\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for different simulation parameters ($DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).  Values represent averages from multiple simulations with common parameters (e.g., row one is a summary of all simulations for which the diel \ac{DO} component was zero).\label{tab:dtd_perf1}} 
\begin{center}
\begin{tabular}{llllllclllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
\cline{2-6} \cline{8-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.30&0.51&0.82&1.00&&0.00&0.68&1.10&1.97&2.39\tabularnewline
~~1&-0.28&0.38&0.59&0.88&1.00&&0.00&0.59&1.07&1.96&2.40\tabularnewline
~~2&-0.39&0.46&0.63&0.90&1.00&&0.00&0.62&1.10&1.97&2.40\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
~~0& 0.00&0.27&0.58&0.93&1.00&&0.00&0.34&1.00&1.96&2.12\tabularnewline
~~1&-0.78&0.37&0.58&0.83&1.00&&0.00&0.63&1.09&1.98&2.12\tabularnewline
~~2&-0.78&0.47&0.61&0.82&1.00&&0.00&0.98&1.34&1.99&2.40\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.34&0.57&0.86&1.00&&0.00&0.63&1.06&1.96&2.40\tabularnewline
~~1&-0.78&0.37&0.59&0.85&1.00&&0.00&0.63&1.06&1.97&2.40\tabularnewline
~~2&-0.78&0.41&0.61&0.85&1.00&&0.00&0.63&1.11&1.98&2.40\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.31&0.82&0.98&1.00&&0.00&0.29&0.76&1.50&2.40\tabularnewline
~~1& 0.05&0.37&0.58&0.81&0.99&&0.07&0.98&1.05&1.49&2.39\tabularnewline
~~2& 0.05&0.40&0.58&0.70&0.99&&0.15&1.06&1.96&2.01&2.40\tabularnewline
\hline
\end{tabular}\end{center}

\end{table}

% summary of simulation performance for detided and biological, window widths
%latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,     caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),     cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,     colheads = rep(c("Min", "25\\textsuperscript{th}", "Median",         "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2")%
\begin{table}[!tbp]
\caption{Summary (range, median, quartiles) of correlations and error estimates comparing filtered and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).  Values represent averages from multiple simulations with common window values (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}} 
\begin{center}
\begin{tabular}{llllllclllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Window}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
\cline{2-6} \cline{8-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Median}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
\hline
{\bfseries Days}&&&&&&&&&&&\tabularnewline
~~1&-0.78&0.63&0.89&0.97&1.00&&0.00&0.28&0.59&1.04&2.12\tabularnewline
~~3&-0.07&0.40&0.59&0.75&1.00&&0.00&0.99&1.08&1.28&2.08\tabularnewline
~~6& 0.00&0.26&0.40&0.58&1.00&&0.00&1.95&1.98&2.05&2.40\tabularnewline
\hline
{\bfseries Hours}&&&&&&&&&&&\tabularnewline
~~1&-0.78&0.36&0.58&0.82&1.00&&0.00&0.63&1.11&1.96&2.40\tabularnewline
~~3& 0.00&0.40&0.60&0.87&1.00&&0.00&0.58&1.07&1.97&2.36\tabularnewline
~~6& 0.03&0.37&0.59&0.85&1.00&&0.00&0.64&1.10&1.98&2.40\tabularnewline
\hline
{\bfseries Tide}&&&&&&&&&&&\tabularnewline
~~0.25& 0.00&0.42&0.63&0.91&1.00&&0.00&0.51&1.04&1.97&2.21\tabularnewline
~~0.5& 0.06&0.43&0.62&0.88&1.00&&0.00&0.61&1.09&1.97&2.27\tabularnewline
~~1&-0.78&0.30&0.51&0.79&1.00&&0.00&0.73&1.20&1.97&2.40\tabularnewline
\hline
\end{tabular}\end{center}

\end{table}

% descriptive table of case studies
%latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,     caption = cap.val, caption.loc = "top", cgroup = c("Tidal amplitude",         "Water quality", "Metabolism\\textsuperscript{\\textit{a}}"),     n.cgroup = c(4, 4, 3), rowname = rows, colheads = c("O1",         "P1", "M2", "S2", "DO", "Chl", "Sal", "Temp", "Pg", "Rt",         "NEM"), label = "tab:case_att")%
\begin{table}[!tbp]
\caption{Summary statistics of tidal component amplitudes (m), selected water quality parameters (\ac{DO} mg L$^{-1}$, chlorophyll-a $\mu$g L$^{-1}$, salinity psu, water temperature $^{\circ}$C)  and metabolism estimates (gross production, respiration, and net ecosystem metabolism as g m$^{-2}$ d$^{-1}$) for each case study.  Tidal components are principal lunar semidiurnal (O1, frequency 25.82 hours), solar diurnal (P1, 24.07 hours), lunar semidiurnal (M2, 12.42 hours), and solar semidiurnal (S2, 12 hours) estimated from harmonic regressions of tidal height (\texttt{oce} package in R, \citealt{Foreman89}, \citetalias{RDCT14}).  Water quality data are averages for the entire period of record for each site.  Metabolism estimates are means of daily integrated values.\label{tab:case_att}} 
\begin{center}
\begin{tabular}{lllllcllllclll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{4}{c}{\bfseries Tidal amplitude}&\multicolumn{1}{c}{\bfseries }&\multicolumn{4}{c}{\bfseries Water quality}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Metabolism\textsuperscript{\textit{a}}}\tabularnewline
\cline{2-5} \cline{7-10} \cline{12-14}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{O1}&\multicolumn{1}{c}{P1}&\multicolumn{1}{c}{M2}&\multicolumn{1}{c}{S2}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Chl}&\multicolumn{1}{c}{Sal}&\multicolumn{1}{c}{Temp}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Pg}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
ELKVM&0.24&0.12&0.48&0.13&&7.87&3.87&32.43&13.78&&8.14&-8.19&-0.05\tabularnewline
PDBBY&0.46&0.23&0.63&0.15&&8.97&2.24&29.17&10.44&&5.95&-5.90& 0.05\tabularnewline
RKBMB&0.13&0.04&0.36&0.10&&4.48&4.50&30.53&25.85&&3.02&-3.62&-0.60\tabularnewline
SAPDC&0.10&0.02&0.54&0.07&&4.96&5.98&27.30&21.77&&4.89&-6.04&-1.16\tabularnewline
\hline
\end{tabular}\end{center}

\footnotesize\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

% correlations with tide before/after wtreg
%latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),     n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,     colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",         "NEM"), caption.loc = "top", rowname = rep(c("Observed",         "Filtered"), 4), label = "tab:cor_res")%
\begin{table}[!tbp]
\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after (filtered) filtering with weighted regression.  Values are averages of monthly correlations.  \ac{DO} values are correlated with predicted tidal height at each observation, whereas metabolism estimates are correlated with mean tidal height change between observations during day, night, or total day periods for production, respiration, and net metabolism, respectively.\label{tab:cor_res}} 
\begin{center}
\begin{tabular}{lllll}
\hline\hline
\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
{\bfseries ELKVM}&&&&\tabularnewline
~~Observed& 0.44& 0.43& 0.43& 0.33\tabularnewline
~~Filtered&-0.04& 0.04&-0.01& 0.09\tabularnewline
\hline
{\bfseries PDBBY}&&&&\tabularnewline
~~Observed&-0.49&-0.11&-0.29&-0.28\tabularnewline
~~Filtered& 0.01&-0.05& 0.00&-0.33\tabularnewline
\hline
{\bfseries RKBMB}&&&&\tabularnewline
~~Observed& 0.45& 0.26& 0.34& 0.30\tabularnewline
~~Filtered& 0.02&-0.04& 0.03& 0.10\tabularnewline
\hline
{\bfseries SAPDC}&&&&\tabularnewline
~~Observed& 0.62& 0.47& 0.64& 0.43\tabularnewline
~~Filtered& 0.00&-0.04& 0.07&-0.04\tabularnewline
\hline
\end{tabular}\end{center}

\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
